---
title: "Prostar navigation tutorial"
subtitle: "or how to use workflow UI"
author: "Samuel Wieczorek"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  BiocStyle::html_document:
    toc: true
    theme: united
    keep_md: true
    papersize: a4
vignette: >
  %\VignetteIndexEntry{Prostar core}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      comment = "#>")
```


```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```


# Introduction

This vignette aims to describe how to use the workflow ui to process a dataset. The goal is not to describe the different data processing tools available in prostar but to explain the behaviour of the user interface of what we call 'Prostar core'.

As explained in xxx, Prostar implements two different levels of data processing tools:
* the upper level correspond to a 'pipeline' and it deals with the type of data and the level of analysis (peptide, protein, peptidomic, etc..). The description of that can be found here (xxx).
* the xxx level correspond to the data processing tools (Filtering, Normalization, etc..)



Color code :
* Grey with solid line: 
* Grey with dashed line:
* Green: the step is validated (the result of the process has been saved)
* Red: the step is mandatory. All further steps are disabled until the user validate the mandatory one


Plain green circle indicates a validated step.
When the colors are transparent, it means that the corresponding step is disabled: all inputs and the reset button are disabled.

By default, when no dataset has been loaded yet, all the steps are disabled