---
title: "Protein workflow with Prostar2"
author: 
- name: Manon Gaudin
- name: Samuel Wieczorek
- name: Thomas Burger
package: Prostar2
date: "`r Sys.Date()`"
output:
    BiocStyle::html_document:
        self_contained: false
        highlight: tango
        toc_float: true
vignette: >
    %\VignetteIndexEntry{Protein workflow with Prostar2}
    %\VignetteEngine{knitr::rmarkdown}
    %%\VignetteKeywords{Mass Spectrometry, Quantitative}
    %\VignetteEncoding{UTF-8}
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

```{r setups, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(123)
```



# Introduction

`Prostar2` is a package for analyzing quantitative data from label-free bottom-up proteomics experiments. It uses `MultiAssayExperiment`, `SummarizedExperiment` and `QFeatures` data structures.

Starting from a table of abundance values and associated metdata, `DaparToolshed` proposes functions to build a complete statistical analysis workflow up to the selection of differentially expressed proteins in the contratsted conditions. `DaparToolshed` makes it possible to import data at different levels: precursor, peptide or protein, each requiring specific processing.

This vignette describes a complete precursor- or peptide-level workflow.

This package is used by the `Prostar2` package, which provides a user-friendly graphical interfaces, so that programming skill is not required to achieve the data analysis.


```{r setup, message = FALSE}
library(Prostar2)
```




# Launching single process




```{r launchProcess, message = FALSE, eval = FALSE}
library(Prostar2)

Prostar2("PipelineProtein_Filtering")
Prostar2("PipelineProtein_Normalization")
Prostar2("PipelineProtein_Imputation")
Prostar2("PipelineProtein_HypothesisTest")
```


Lorsque l'on lance un seul process, une interface dans la partie Description permet de charger un jeu de données.


# Launching a pipeline (a serie of processes)

```{r launchPipeline, message = FALSE, eval = FALSE}
library(Prostar2)
Prostar2("PipelineProtein")
```


Lorsque l'on lance un pipeline, l'interface est un peu plus complète que pour un process seul. Dans cette interface, il y a un menu rétractable sur la gauche qui contient plusieurs items :
* Dataset manager
  * Open a dataset
  * Import (convert a csv or Excel file to a *qf file)
  * Save
* Run workflow
  * xxxxx
  * xxxx
  * xxxx
  


